---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    fig_caption: yes
    keep_md: yes
---


## Loading and preprocessing the data

 
   
```{r, echo=TRUE}
 temp <- tempfile()
    download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp, mode="wb")
    unzip(temp, "activity.csv")
    activity_data <- read.table("activity.csv", sep=",", header=T)
```


## What is mean total number of steps taken per day?

   
    
### Histogram of the total number of steps taken each day


```{r, echo=TRUE}
 agg_data<-aggregate(activity_data$steps, list(activity_data$date),sum)
    names(agg_data)<- c("date","total_steps")
    agg_data$date<-as.POSIXct(agg_data$date)
    plot(agg_data$date,agg_data$total_steps,type = 'h', lwd=10,col="purple",xlab='Date', ylab='Total number of     steps', main = 'Total steps taken each day')
    
```

   
   
##R code to calculate the mean and median of total no. of steps taken each day

    
   
```{r, echo=TRUE}
  mean_data<-aggregate(activity_data$steps,list(activity_data$date),mean)
    names(mean_data)<- c("date","mean_value")
    mean_data
    total_mean<-mean(agg_data$total_steps,na.rm = TRUE)
    total_mean
    total_median<- median(agg_data$total_steps,na.rm = TRUE)
    total_median
```


## What is the average daily activity pattern?

##R code to plot a time series for 5-minute interval
  
```{r, echo=TRUE}
      activity_data_not_na<-activity_data[!(is.na(activity_data$steps)),]
      interval_data<-aggregate(activity_data_not_na$steps,list(activity_data_not_na$interval),mean)
      names(interval_data)<-c("interval","avg_steps")
      plot(interval_data$interval,interval_data$avg_steps,type = 'l',xlab =  "Interval",ylab = "Average Steps",       main = "Average Steps for 5-minute Interval", col="orange")
      max_interval<-interval_data[which.max(interval_data$avg_steps),]
      max_interval
```  
  
## Imputing missing values
```{r, echo=TRUE}
    total_missing_rows<-sum(is.na(activity_data))
    total_missing_rows
    activity_data2<-activity_data
    missing_data<-is.na(activity_data2$steps)
    mean_value<-tapply(activity_data_not_na$steps,activity_data_not_na$interval,mean,na.rm=TRUE,simplify = TRUE)
    activity_data2$steps[missing_data]<-mean_value[as.character(activity_data2$interval[missing_data])]
    sum(is.na(activity_data2))
```
##R code to plot histogram on new dataset
```{r, echo=TRUE}
    agg_data2<-aggregate(activity_data2$steps, list(activity_data2$date),sum)
    names(agg_data2)<- c("date","total_steps")
    agg_data2$date<-as.POSIXct(agg_data2$date)
    plot(agg_data2$date,agg_data2$total_steps,type = 'h', lwd=10,col="purple",xlab='Date', ylab='Total number of steps', main = 'Total steps taken each day')
```
##R code to calculate the mean and median of total no. of steps taken each day on new dataset
```{r, echo=TRUE}
    mean_data2<-aggregate(activity_data2$steps,list(activity_data2$date),mean)
    names(mean_data2)<- c("date","mean_value")
    mean_data2
    total_mean2<-mean(agg_data2$total_steps,na.rm = TRUE)
    total_mean2
    total_median2<- median(agg_data2$total_steps,na.rm = TRUE)
    total_median2
```
##The new and old mean remains same but the new median incresed by 1.19
```{r, echo=TRUE}
    paste("Old Mean: ", total_mean )
    paste("New Mean: ", total_mean2)
    paste("Old Median: ", total_median)
    paste("New Median: ", total_median2)
```
##R code to add a new variable for weekday
```{r, echo=TRUE}
    library(ggplot2)
    activity_data2$date<-as.POSIXct(activity_data2$date)
    activity_data2$day <- weekdays(activity_data2$date)
    activity_data2$day_type<-ifelse(activity_data2$day == 'Sunday' | activity_data2$day =='Saturday', 'Weekend','Weekday')
    activity_data2$day_type<-as.factor(activity_data2$day_type)
    interval_data2<-aggregate(activity_data2$steps,list(activity_data2$interval,activity_data2$day_type),mean)
    names(interval_data2)<-c("interval","day_type","avg_steps")
    ggplot(interval_data2,aes(interval,avg_steps))+geom_line(col="navy blue")+facet_grid(.~day_type)+labs(x="Interval",y="Avg Steps",title="Avg steps on Weekday/weekend")
```    
